---
title: "Decathlon analysis"
output: html_notebook
---
Libraries I will be using:

```{r}
library(tidyverse)
library(readr)
library(here)
library(stringr)
```

Reading the data

```{r}
decathlon_clean_data <- read_csv(here("clean_data/decathlon_data_clean.csv"))

decathlon_clean_data

```

## Who had the longest long jump seen in the data?

```{r}
longest_jump <- decathlon_clean_data %>% 
  filter(sport_rank_points == "long_jump") %>% #filtering to just result for long_jump
  arrange(desc(score)) %>% 
  top_n(1)

longest_jump
```

Clay competing in the Olympic Games has the longest jump in the data with a score of 7.96

##What was the average 100m time in each competition?

```{r}
decastar_100_meters <- decathlon_clean_data %>% 
  filter(sport_rank_points == "100_meters", competition == "decastar")   

#creating a new variable which includes just 100 meters result from the Decastar competition 
  
decastar_100_meters_avg <- mean(decastar_100_meters$score, na.rm = TRUE)

#calculating the average

decastar_100_meters_avg
```
Average 100m time for Decastar competion is 11.17538

```{r}
olympics_100_meters <- decathlon_clean_data %>% 
  filter(sport_rank_points == "100_meters", competition == "olympic_games")   

#creating a new variable which includes just 100 meters result from the Olympic Games competition 
  
olympics_100_meters_avg <- mean(olympics_100_meters$score, na.rm = TRUE)

#calculating the average

olympics_100_meters_avg

```

Average 100m time for Olympic Games competition is 10.91571

##Who had the highest total points across both competitions?

```{r}
highest_total_points <- decathlon_clean_data %>%
  filter(sport_rank_points == "points") %>% #filtering the total point for each athlete 

highest_total_points

```

Sebrle has the highest high score across both competitioons. 

##What was the shot-put scores for the top three competitors in each competition?

Top three shot-put score for each competition:

```{r}
shot_put_scores_decastar <- decathlon_clean_data %>%
  filter(sport_rank_points == "shot_put", competition == "decastar") %>% 
  arrange(desc(score)) %>% 
  top_n(3)
  
shot_put_scores_decastar
```

```{r}
shot_put_scores_olympic_games <- decathlon_clean_data %>%
  filter(sport_rank_points == "shot_put", competition == "olympic_games") %>% 
  arrange(desc(score)) %>% 
  top_n(3)
  
shot_put_scores_olympic_games
```

# Top three competitors in each competition shot_put score

```{r}
athletes_points <- decathlon_clean_data %>% #creating a data set with athletes points
  filter(sport_rank_points == "points") %>% 
  rename(points = score) %>% 
  select("athlete","points")
  
shot_put_score <- decathlon_clean_data %>% #creating a data set with the athletes shot-put scores
  filter(sport_rank_points == "shot_put") %>% 
  rename(score_shot_put = score) %>% 
  select("athlete", "competition", "score_shot_put")
  
shot_put_joined <-full_join(shot_put_score, athletes_points) #joining the two new dataset

#getting the top 3 for each category:

top_three_decastar <- shot_put_joined %>% 
  filter(competition == "decastar") %>% 
  arrange(desc(points)) %>% 
  top_n(3)

top_three_decastar

top_three_olympics <- shot_put_joined %>% 
  filter(competition == "olympic_games") %>% 
  arrange(desc(points)) %>% 
  top_n(3)

top_three_olympics

#it would have be best if I haven't included the points in the sport_rank_points column during the cleaning

```

Decastart Top 3 shot-put score :

Sebrle	14.83		
Clay	14.26		
Karpov	14.77	

Olympic Top 3 shot-put score

Sebrle 16.36
Macey	15.73	 	
Clay 15.23	

##What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?

```{r}
decathlon_400_m <- decathlon_clean_data %>% #creating a data set with 400 meters details 
  filter(sport_rank_points == "400_meters")

joined_400_m <- inner_join(decathlon_400_m, athletes_points) #joining the new data set with the athletes_points

#getting the averages: 

avg_points_less_50 <- joined_400_m %>% 
  filter(score <= 50) %>% 
  summarise(mean(points))

avg_points_less_50

avg_points_more_50 <- joined_400_m %>% 
  filter(score >= 50) %>% 
  summarise(mean(points))

avg_points_more_50

```

Average points less than 50 is 8146.087	
Average points more than 50 is 7766.385		